Taking uncertainty in word embedding bias estimation seriously: a bayesian approach

### 2023-05-25 paper/paperDraft.tex
revised visualizations, fixed intro

\usepackage{sectsty}
\sectionfont{\fontsize{11}{11}\selectfont}

@@

Natural language processing (NLP) models are used to perform various
language-related tasks such as providing email filters, intelligent
assistants, search results, translations, text analysis and so on. Such
tools require input words to be represented by vectors of real numbers.
This can usually be done with word embeddings. A common way to construct
an embedding is to use a large natural language corpus to train a neural
network to assign vectors to words that optimize for co-occurrence
prediction accuracy. The vectors can then be compared in terms of their
similarity -- the usual measure is cosine similarity -- and the results
of such comparisons can be used in downstream tasks. Roughly speaking,
cosine similarity is an imperfect mathematical proxy for semantic
similarity {[}14{]}.

It has been suggested {[}1,4,5,8,11,12{]} that such models can learn
implicit biases that reflect harmful stereotypical thinking---for
example, the (vector corresponding to the) word \textit{she} might be
much closer in the vector space to the word \textit{cooking} than the
word \textit{he}. Such phenomena may be undesirable at least in some
downstream tasks, such as web search, recommendations, and so on. To
investigate such issues, several measures of bias in word embeddings
have been formulated and applied. Our goal is to use two prominent
examples of such measures to argue that this approach oversimplifies the
situation and to develop a Bayesian alternative.

@@

Another response might be to insist that there is no bias introduced by
the use of machine learning methods here, since the algorithm is simply
learning to correctly predict co-occurrences based on what ``reality''
looks like. However, this objection overlooks the fact that we, humans,
are the ones who construct this linguistic reality, is shaped in part by
the natural language processing tools we use on a massive scale. Sure,
if there is unfairness and our goal is to diagnose it, we should do
complete justice to learning it in the model used in a study thereof.
One example of this approach is {[}3{]}. However, if our goal is to
develop downstream tools that perform tasks that we care about without
further perpetuating or exacerbating harmful stereotypes, we still have
good reasons to try to minimize the negative impact, at least as long as
this does not damage the overall performance of the tool. Moreover, it
is often not the case that the corpora mirror reality---to give a
trivial example, heads are spoken of more often than kidneys, but this
does not mean that kidneys occur much less often in reality than heads.
To give a more relevant example, the disproportionate association of
female words with female occupations in a corpus actually greatly
exaggerates the actual lower disproportion in the real distribution of
occupations. {[}6{]}.

In what follows, we focus on two popular measures of bias applicable to
many existing word embeddings (
\emph{GoogleNews},\footnote{GoogleNews-vectors-negative300, available at  \url{https://github.com/mmihaltz/word2vec-GoogleNews-vectors}.}
\emph{Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.}
and
\emph{Reddit Corpus}\footnote{The Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}):
\emph{Word Embedding Association Test} (\textsf{WEAT}) {[}8{]}, and
\emph{Mean Average Cosine Distance} (\textsf{MAC}) {[}12{]}. We first
explain how these measures are supposed to work. Then we argue that they
are problematic for various reasons, a key one being that by
pre-averaging data they manufacture false confidence, which we
illustrate in terms of simulations showing that the measures often
suggest the existence of bias even if by design it is non-existent in a
simulated dataset.

We propose to replace them with a Bayesian data analysis (hierarchical
models), which not only provides more modest and realistic assessment of
the uncertainty involved, but also allows for inspection at various
levels of granularity. Once we introduce the method, we apply it to
multiple word embeddings and results of supposed debiasing, putting
forward some general observations that are not exactly in line with the
usual picture painted in terms of \textsf{WEAT} or \textsf{MAC} (and the
problem generalizes to any approach that focuses on chasing a single
numeric metric): the word list sizes and sample sizes used in the
studies are usually small,\footnote{Depending on a list for {[}8{]} the
  range for protected words is between 13 and 100, and for attributes
  between 16 and 25; for {[}12{]} the range for protected words is
  between 14 and 18, and for attributes between 11 and 25.} posterior
density intervals are fairly wide, often the differences between
associated, different and neutral human predicates, are not very
impressive. Also, a preliminary inspection suggests that the
desirability of changes obtained by the usual debiasing methods is
debatable. The tools that we have proposed, however, allow for a more
fine-grained and multi-level evaluation of bias and debiasing in
language models without abandoning modesty about the uncertainties
involved.

@@

Conceptually the idea is that if a particular harmful stereotype is
learned in a particular embedding, then certain groups of words will be
systematically closer to (or further from) each other. This gives rise
to the idea of protected groups---for example, in guiding online search
completion or recommendation, female words might require protection in
that they should not be systematically closer to stereotypically female
job names, and male words require protection in that they should not be
systematically closer to toxic masculinity stereotypes, such as
``tough'', ``never complaining'' or ``macho''.\footnote{However, for
  some research-related purposes, such as the study of stereotypes
  across history {[}3{]}, embeddings which do not protect certain
  classes may also be useful.}

The key role in the measures to be discussed is played by the notion of
cosine distance (or, symmetrically, by cosine similarity). These are
defined as follows:\footnote{Here, ``\(-\)'' stands for point-wise
  difference, ``\(\cdot\)'' stands for the dot product operation, and
  \(\lVert a\rVert = \sqrt{(a \cdot a)}\).} \begin{align} \tag{Sim}
\mathsf{cosineSimilarity}(A,B) & = \frac{A \cdot B}{\lVert  A \rVert \,\lVert B \rVert}
\\
\tag{Distance}
\mathsf{cosineDistance}(A,B) &  = 1 - \mathsf{cosineSimilarity}(A,B).
\end{align} Note that this terminology is slightly misleading, as
mathematically cosine distance is not a distance measure, because it
does not satisfy the triangle inequality, as generally
\(\mathsf{cosineDistance}(A,C) \not \leq \mathsf{cosineDistance}(A,B)+ \mathsf{cosineDistance}(B,C)\).
We will keep using this mainstream terminology.

@@

\hypertarget{methodological-problems-with-cosine-based-bias-metrics}{%
\section{Methodological problems with cosine-based bias
metrics}\label{methodological-problems-with-cosine-based-bias-metrics}}

@@

In Figure \ref{fig:empirical} we look at the empirical distributions,
while paying attention to such divisions. The horizontal lines represent
the values of 1-\textsf{MAC} (that is, we now talk in terms of cosine
similarity rather than cosine distance; this is a trivial linear
transformation) that the authors considered indicative of bias for
stereotypes corresponding to given word lists. For instance, in
religion, \textsf{MAC} was .859 and considered a sign of bias, so we
plot \(0\pm (1-.859)\approx .14\) lines around similarity = 0 (that is,
distance = 1). Notice that most distributions are quite wide, and the
proportions of even neutral or human neutral words with similarities
higher than the value of 1-\textsf{MAC} deserving debiasing according to
the authors is quite high.

@@

\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8b-1} \end{center}

\caption{Empirical distributions of cosine similarities (1-distances) for three word lists used in  the original paper.  }

\label{fig:empirical}
\end{figure}

@@

\hypertarget{pre-averaging-bootstrapping-and-testing-with-the-null-model}{%
\section{Pre-averaging, bootstrapping and testing with the null
model}\label{pre-averaging-bootstrapping-and-testing-with-the-null-model}}

The approaches we have been describing use means of mean average cosine
similarities to measure similarity between protected words and
attributes coming from harmful stereotypes. However, a closer inspection
look at the individual values that are taken for the calculations, it
turns out that there are quite a few outliers and surprisingly
dissimilar words. This problem becomes transparent when we examine the
visualizations of the individual cosine distances, following the idea
that one of the first steps in understanding data is to look at it.

@@

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/visMuslimd-1} \end{center}

@@

\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:ourDistancesPlot2-1} \end{center}

\caption{Distributions of test statistics and effect sizes based on 10k simulations on the assumption of a null model in which all distances come from normal distribution with $\mu =0, \sigma = .08$.}
\label{fig:ourDistances}
\end{figure}

@@

\hypertarget{introduction-and-model-definition}{%
\subsection{Introduction and model
definition}\label{introduction-and-model-definition}}

@@

@@

\vspace{1mm}
\footnotesize

@@

@@

These visualizations can be also handy when it comes to the
investigation of the effect that debiasing has on the embedding space.
Below one may see chosen pair depicting the difference in means with
89\% highest posterior density intervals before and after applying
debiasing.

@@

\begin{center}\includegraphics[width=0.85\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosinePair2b-1} \includegraphics[width=0.85\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosinePair2b-2} \end{center}
\caption{Means with 89\% highest posterior density intervals for gender before and after debiasing.}
\label{fig:empiricalDebiasedPair}
\end{figure}

In Figure \ref{fig:empiricalDebiased} we look at the empirical
distributions for the debiased embeddings. Comparing the results to the
original embedding, one may notice that for the Religion group the
neutral and human distribution has changed slightly. Before within the
``correct'' cosine similarity boundaries there were 56\% of neutral and
55\% of human word lists. After the debiasing the values changed to 59\%
(for neutral) and 59\% (for human). The different and associated word
lists were more influenced. The general shape of both distributions is
less stretched. Before debiasing 43\% of the different word lists and
35\% of the associated word lists were within the accepted boundaries.
After the embedding manipulation the percentage has increased for both
lists to 63\%. Visualization for Gender group illustrates almost no
change for the neutral and human word lists before and after debiasing.
The values for different and associated word lists are also barely
impacted by the embedding modification. In the Race group, the
percentage within the boundaries for neutral and associated word lists
have increased. The opposite happened for human and different word
lists, where the percentage of ``correct'' cosine similarity dropped
from 67\% to 55\% (human) and from 39\% to 36\% (different).

@@

\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine-1} \end{center}
\caption{Empirical distributions of debiased cosine similarities (1-distances) for three word lists used in  the original paper.}
\label{fig:empiricalDebiased}
\end{figure}

@@

@@

\begin{center}\includegraphics{paperDraft_files/figure-latex/weatjoint2r-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/weat7glovf-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat7Reddit-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Google-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Glove-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Reddit-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGoogle-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGlove-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionReddit-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGoogle-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGlove-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceReddit-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGoogle-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGlove-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderReddit-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedReligionReddit-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedRaceReddit-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedGenderReddit-1} \end{center}

\hypertarget{lists-for-manzini2019blacktocriminal}{%
\subsubsection{Lists for
{[}12{]}:}\label{lists-for-manzini2019blacktocriminal}}

\label{appendix:manzini_word_lists} The lists are available here:

\hypertarget{our-custom-word-lists}{%
\subsubsection{Our custom word lists}\label{our-custom-word-lists}}

@@

>>>>

% \usepackage{sectsty}
% \sectionfont{\fontsize{12}{11}\selectfont}
% \subsectionfont{\fontsize{11}{11}\selectfont}

@@

It has been suggested\footnote{See for instance {[}1,4,5,8,11,12{]}.}
that language models can learn implicit biases that reflect harmful
stereotypical thinking---for example, the (vector corresponding to the)
word \textit{she} might be much closer in the vector space to the word
\textit{cooking} than the word \textit{he}. Such phenomena are
undesirable at least in some downstream tasks, such as web search,
recommendations, and so on. To investigate such issues, several measures
of bias in word embeddings have been formulated and applied. Our goal is
to use two prominent examples of such measures to argue that this
approach oversimplifies the situation and to develop a Bayesian
alternative.

A common approach in natural language processing is to represent words
by vectors of real numbers---such representations are called
\emph{embeddings}. One way to construct an embedding---we will focus our
attention on non-contextual language models\footnote{One example of a
  contextualized representation is BERT. Another is GPT.}---is to use a
large corpus to train a neural network to assign vectors to words in a
way that optimizes for co-occurrence prediction accuracy. Such vectors
can then be compared in terms of their similarity---the usual measure is
cosine similarity---and the results of such comparisons can be used in
downstream tasks. Roughly speaking, cosine similarity is an imperfect
mathematical proxy for semantic similarity {[}14{]}.

@@

Another response might be to insist that there is no bias introduced by
the use of machine learning methods here, since the algorithm is simply
learning to correctly predict co-occurrences based on what ``reality''
looks like. However, this objection overlooks the fact that we, humans,
are the ones who construct this linguistic reality, which is shaped in
part by the natural language processing tools we use on a massive scale.
Sure, if there is unfairness and our goal is to diagnose it, we should
do complete justice to learning it in the model used to study it. One
example of this approach is {[}3{]}, where the authors use language
models to study the shape of certain biases across a century.

However, if our goal is to develop downstream tools that perform tasks
that we care about without further perpetuating or exacerbating harmful
stereotypes, we still have good reasons to try to minimize the negative
impact. Moreover, it is often not the case that the corpora mirror
reality---to give a trivial example, heads are spoken of more often than
kidneys, but this does not mean that kidneys occur much less often in
reality than heads. To give a more relevant example, the
disproportionate association of female words with female occupations in
a corpus actually greatly exaggerates the actual lower disproportion in
the real distribution of occupations {[}6{]}.

In what follows, we focus on two popular measures of bias applicable to
many existing word embeddings, such as
\emph{GoogleNews},\footnote{GoogleNews-vectors-negative300, available at  \url{https://github.com/mmihaltz/word2vec-GoogleNews-vectors}.}
\emph{Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.}
and
\emph{Reddit Corpus}\footnote{Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}:
\emph{Word Embedding Association Test} (\textsf{WEAT}) {[}8{]}, and
\emph{Mean Average Cosine Distance} (\textsf{MAC}) {[}12{]}. We first
explain how these measures are supposed to work. Then we argue that they
are problematic for various reasons---the key one being that by
pre-averaging data they manufacture false confidence, which we
illustrate in terms of simulations showing that the measures often
suggest the existence of bias even if by design it is non-existent in a
simulated dataset.

We propose to replace them with a Bayesian data analysis, which not only
provides more modest and realistic assessment of the uncertainty
involved, but in which hierarchical models allow for inspection at
various levels of granularity. Once we introduce the method, we apply it
to multiple word embeddings and results of supposed debiasing, putting
forward some general observations that are not exactly in line with the
usual picture painted in terms of \textsf{WEAT} or \textsf{MAC}.

Most of the problems that we point out generalize to any existing
approach that focuses on chasing a single numeric metric of bias. (1)
They treat the results of pre-averaging as raw data in statistical
significance tests, which in this context is bound to overestimate
significance. We show similar results can easily be obtained when
sampling from null models with no bias. (2) The word list sizes and
sample sizes used in the studies are usually small,\footnote{Depending
  on a list for {[}8{]} the range for protected words is between 13 and
  100, and for attributes between 16 and 25; for {[}12{]} the range for
  protected words is between 14 and 18, and for attributes between 11
  and 25.} (3) Many studies do not use any control predicates, such as
random neutral words or neutral human predicates for comparison.

On the constructive side, we develop and deploy our method, and the
results are, roughly, as follows. (A) Posterior density intervals are
fairly wide and the average differences between associated, different
and neutral human predicates, are not very large. (B) A preliminary
inspection suggests that the desirability of changes obtained by the
usual debiasing methods is debatable.

In Section \ref{sec:two} we describe the two key measures discussed in
this paper, \textsf{WEAT} and \textsf{MAC}, explaining how they are
calculated and how they are supposed to work. In Section
\ref{sec:challenges} we first argue in Subsection
\ref{subsec:interpretability}, that it is far from clear how results
given in terms of \textsf{WEAT} or \textsf{MAC} are to be interpreted.
Second, in Subsection \ref{subsec:problems} we explain the statistical
problems that arise when one uses pre-averaged data in such contexts, as
these measures do. In Section \ref{sec:bayesian} we explain the
alternative Bayesian approach that we propose. In Section
\ref{sec:results} we elaborate on the results that it leads to,
including a somewhat skeptical view of the efficiency of debiasing
methods, discussed in Subsection \ref{subsec:rethinking}. Finally, in
Section \ref{sec:related} we spend some time placing our results in the
ongoing discussions.\footnote{\textbf{Disclaimer:} throughout the paper
  we will be mentioning and using word lists and stereotypes we did not
  formulate, which does not mean we condone any judgment made therein or
  underlying a given word selection. For instance, the Gender dataset
  does not recognize non-binary categories, and yet we use it without
  claiming that such categories should be ignored.}

@@

\label{sec:two}

Conceptually, the idea is that if a particular harmful stereotype is
learned in a particular embedding, then certain groups of words will be
systematically closer to (or further from) each other. This gives rise
to the idea of protected groups---for example, in guiding online search
completion or recommendation, female words might require protection in
that they should not be systematically closer to stereotypically female
job names, such as ``nurse'', ``librarian'', ``waitress'', and male
words require protection in that they should not be systematically
closer to toxic masculinity stereotypes, such as ``tough'', ``never
complaining'' or ``macho''.\footnote{However, for some research-related
  purposes, such as the study of stereotypes across history {[}3{]},
  embeddings which do not protect certain classes may also be useful.}

The key role in the measures to be discussed is played by the notion of
cosine distance (or, symmetrically, by cosine similarity). These are
defined as follows:\footnote{Here, ``\(-\)'' stands for point-wise
  difference, ``\(\cdot\)'' stands for the dot product operation, and
  \(\lVert a\rVert = \sqrt{(a \cdot a)}\).}
\(^{\!\!\! , \,}\)\footnote{Note that this terminology is slightly
  misleading, as mathematically cosine distance is not a distance
  measure, because it does not satisfy the triangle inequality, as
  generally
  \(\mathsf{cosineDistance}(A,C) \not \leq \mathsf{cosineDistance}(A,B)+ \mathsf{cosineDistance}(B,C)\).
  We will keep using this mainstream terminology.}
\begin{align} \tag{Sim}
\mathsf{cosineSimilarity}(A,B) & = \frac{A \cdot B}{\lVert  A \rVert \,\lVert B \rVert}
\\
\tag{Distance}
\mathsf{cosineDistance}(A,B) &  = 1 - \mathsf{cosineSimilarity}(A,B).
\end{align}

@@

\hypertarget{challenges-to-cosine-based-bias-metrics}{%
\section{Challenges to cosine-based bias
metrics}\label{challenges-to-cosine-based-bias-metrics}}

\label{sec:challenges}

\hypertarget{interpretability-issues}{%
\subsection{Interpretability issues}\label{interpretability-issues}}

\label{subsec:interpretability}

@@

In Figures (\ref{fig:empirical0}-\ref{fig:empirical3}) we look at the
empirical distributions, while paying attention to such divisions. The
horizontal lines represent the values of 1-\textsf{MAC} (that is, we now
talk in terms of cosine similarity rather than cosine distance; this is
a trivial linear transformation) that the authors considered indicative
of bias for stereotypes corresponding to given word lists. For instance,
in religion, \textsf{MAC} was .859 and considered a sign of bias, so we
plot \(0\pm (1-.859)\approx .14\) lines around similarity = 0 (that is,
distance = 1). Notice that most distributions are quite wide, and the
proportions of even neutral or human neutral words with similarities
higher than the value of 1-\textsf{MAC} deserving debiasing according to
the authors is quite high.

@@

\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8d-1} \end{center}

\caption{Empirical distributions of cosine similarities  for the Religion word list  used in  the original paper. }

\label{fig:empirical0}
\end{figure}

\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8df-1} \end{center}

\caption{Empirical distributions of cosine similarities for the Gender word list used in  the original paper.  }

\label{fig:empirical2}
\end{figure}

\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8dg-1} \end{center}

\caption{Empirical distributions of cosine similarities  for the Race word list used in  the original paper.  }

\label{fig:empirical3}
\end{figure}

@@

\hypertarget{problems-with-pre-averaging}{%
\subsection{Problems with
pre-averaging}\label{problems-with-pre-averaging}}

\label{subsec:problems}

The approaches we have been describing use means of mean average cosine
similarities to measure similarity between protected words and
attributes coming from harmful stereotypes. However, a closer inspection
look at the individual values that are taken for the calculations, it
turns out that there are quite a few outliers and surprisingly
dissimilar words. This problem becomes transparent when we examine the
visualizations of the individual cosine distances, following the idea
that one of the first steps in understanding data is to look at it.
Let's start with inspecting two examples in Figures \ref{fig:muslim} and
\ref{fig:priest}.

@@

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/visMuslimd-1} \end{center}
\caption{Actual distances for the protected word "muslim". The most similar attributes tend to be the ones associated with it stereotypically, but then words associated with other stereotypes come closer than neutral or human predicates. }
\label{fig:muslim}
\end{figure}

@@

\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:ourDistancesPlot2e-1} \end{center}

\caption{Distributions of test statistics and effect sizes based on 10k simulations on the assumption of a null model in which all distances come from normal distribution with $\mu =0, \sigma = .08, n=10k$.}
\label{fig:ourDistances}
\end{figure}

@@

\label{sec:bayesian}

\hypertarget{model-construction}{%
\subsection{Model construction}\label{model-construction}}

\label{subsec:model}

@@

\label{subsec:posterior}

@@

\label{sec:results}

\vspace{1mm}

@@

\label{subsec:observations}

@@

\label{subsec:rethinking}

These visualizations can be also handy when it comes to the
investigation of the effect that debiasing has on the embedding space.
In Figures \ref{fig:empiricalPriorToDebiasing} and
\ref{fig:empiricalDebiased} we see chosen pair depicting the difference
in means with 89\% highest posterior density intervals before and after
applying debiasing.

@@

\begin{center}\includegraphics{paperDraft_files/figure-latex/fig:debiasedCosinePair2e-1} \end{center}
\caption{Mean cosine distances with 89\% highest posterior density intervals for the  gender dataset before debiasing.}
\label{fig:empiricalPriorToDebiasing}
\end{figure}

@@

\begin{center}\includegraphics{paperDraft_files/figure-latex/fig:debiasedCosinePair2_e-1} \end{center}
\caption{Mean with 89\% highest posterior density intervals for gender after debiasing.}
\label{fig:empiricalDebiased}
\end{figure}

In Figures \ref{fig:empiricalDebiased}-\ref{fig:empiricalDebiased3} we
inspect the empirical distributions for the debiased embeddings.
Comparing the results to the original embedding, one may notice that for
the Religion group the neutral and human distribution has changed
slightly. Before within the ``correct'' cosine similarity boundaries
there were 56\% of neutral and 55\% of human word lists. After the
debiasing the values changed to 59\% (for neutral) and 59\% (for human).
The different and associated word lists were more influenced. The
general shape of both distributions is less stretched. Before debiasing
43\% of the different word lists and 35\% of the associated word lists
were within the accepted boundaries. After the embedding manipulation
the percentage has increased for both lists to 63\%. Visualization for
Gender group illustrates almost no change for the neutral and human word
lists before and after debiasing. The values for different and
associated word lists are also barely impacted by the embedding
modification. In the Race group, the percentage within the boundaries
for neutral and associated word lists have increased. The opposite
happened for human and different word lists, where the percentage of
``correct'' cosine similarity dropped from 67\% to 55\% (human) and from
39\% to 36\% (different).

\begin{center}\includegraphics[width=0.95\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine21c-1} \end{center}
\caption{Empirical distributions of cosine similarities before and after  debiasing  for  the Religion word list  used in  the original paper.}
\label{fig:empiricalDebiased}
\end{figure}

\begin{center}\includegraphics[width=0.95\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine22c-1} \end{center}
\caption{Empirical distributions of cosine similarities before and after  debiasing for  the Gender word list used in  the original paper.}
\label{fig:empiricalDebiased2}
\end{figure}

\begin{center}\includegraphics[width=0.95\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine23c-1} \end{center}
\caption{Empirical distributions of cosine similarities before and after  debiasing for the Race word list used in  the original paper.}
\label{fig:empiricalDebiased3}
\end{figure}

@@

\label{sec:related}

@@

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

@@

\label{sec:appendix}

\begin{center}\includegraphics{paperDraft_files/figure-latex/weatjoint2z-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/weat7glovh-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat7Reddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Googlea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Glovea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Reddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGooglea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGlovea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionReddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGooglea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGlovea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceReddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGooglea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGlovea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderReddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedReligionReddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedRaceReddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedGenderReddita-1} \end{center}

\label{appendix:posterior}

\label{appendix:word}

\hypertarget{lists-used-in-previous-research}{%
\subsubsection{Lists used in previous
research}\label{lists-used-in-previous-research}}

\label{appendix:manzini_word_lists} The lists from {[}12{]} are
available here:

\hypertarget{custom-lists-used-in-this-paper}{%
\subsubsection{Custom lists used in this
paper}\label{custom-lists-used-in-this-paper}}
